---
title: "Dynamic Inconsistency in Food Choice"
author: "Brock Wilson"
date: "4/5/2021"
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)
library(pacman)
```

---
#Summary

Field experiment: home grocery delivery program

Experimenters offer temptation foods at time of delivery in exchange for purchased foods

Findings imply that 40% are dynamically inconsistent

Individuals who are inconsistent are the least likely to use commitment device

---
#Introduction

Temptation and self control models are hot in behavioral economics

Dynamic inconsistencies pose issues: individuals want to go to the gym, but never want to go when asked

This paper aims to address a core question: 

### What is the relationship between dynamically inconsistent behaviour and beliefs thereof? 

In other words: Are inconsistent individuals blind to their faults? 

---
#Introduction

Previous literature finds positive to zero correlation between self control problems and awareness

Suggests that those with self control problems should be aware of their problem

Caveat: all of these findings comes from lab experiments (may have experimental bias)

---

#Aside

Commitment Demand = choosing to restrict oneself to the advance bundle

Advance bundle = bundle pre-ordered

Immediate bundle = bundle with temptation goods

---
#Findings

46% of individuals are dynamically inconsistent when commitment is not available

Regularly individuals substitute fruits and veggies for caloric-dense and high fat foods

When commitment is available, 53% of individuals take it up

- Subjects who are dynamically inconsistent are the least likely to take up commitment (44%)

- Subjects who are dynamically consistent are more likely to take up commitment (60%)

Suggests that those with dynamic inconsistent behavior may be the least aware of their behavior

---

#Implications

How should we measure welfare in these situations?

- Utilizing "long run" preferences without temptations?

- Utilizing "short run" preferences with temptations?

- Utilizing unambiguous preferences? (preferences that are consistent across short and long run)

---

#Experimental Setup

Two experiments used (one for findings and one for replication of findings)

---

#Experimental Setup: Chicago - High Level Summary

Subjects were not told they were apart of an experiment

20 foods were offered in the experiment for home delivery

Subjects were asked to rate each food on a 1 to 7 scale

2 week food delivery program


---

#Experimental Setup: Chicago - Details

Subjects were signed up at a table at grocery stores

Recruitment was on a rolling basis

Research assistants were in charge of setting up experiment and delivering foods

Foods were selected in consultation with store managers

Foods were assured to not be bruised or damaged

In return for participating, subjects received $25-20 gift cards

---
#Experimental Setup: Chicago - Food Selection
```{r fig.width=100, fig.height=55,echo=FALSE}
library(png)
library(grid)
img <- readPNG("/Users/Brock/Desktop/Second-Year-PhD/3rd-Quarter/Experimental-Economics/Paper-Presentation/Exp-Paper-Prez/chicagofoods.png")
grid.raster(img)
```
---

#Experimental Setup: Chicago - Timeline





---
#Experimental Setup: Los Angeles - Food Selection
```{r fig.width=100, fig.height=55,echo=FALSE}
library(png)
library(grid)
img <- readPNG("/Users/Brock/Desktop/Second-Year-PhD/3rd-Quarter/Experimental-Economics/Paper-Presentation/Exp-Paper-Prez/lafoods.png")
 grid.raster(img)
```

---

















