---
title: "Dynamic Inconsistency in Food Choice"
author: "Brock Wilson"
date: "4/5/2021"
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)
library(pacman)
```

---
#Summary

Field experiment: home grocery delivery program

Experimenters offer temptation foods at time of delivery in exchange for purchased foods

Findings imply that 40% are dynamically inconsistent

Individuals who are inconsistent are the least likely to use commitment device

---
#Introduction

Temptation and self control models are hot in behavioral economics

Dynamic inconsistencies pose issues: individuals want to go to the gym, but never want to go when asked

This paper aims to address a core question: 

### What is the relationship between dynamically inconsistent behaviour and beliefs thereof? 

In other words: Are inconsistent individuals blind to their faults? 

---
#Introduction

Previous literature finds positive to zero correlation between self control problems and awareness

Suggests that those with self control problems should be aware of their problem

Caveat: all of these findings comes from lab experiments (may have experimental bias)

---

#Aside

Commitment Demand = choosing to restrict oneself to the advance bundle

Advance bundle = bundle pre-ordered

Immediate bundle = bundle with temptation goods

---
#Findings

46% of individuals are dynamically inconsistent when commitment is not available

Regularly individuals substitute fruits and veggies for caloric-dense and high fat foods

When commitment is available, 53% of individuals take it up

- Subjects who are dynamically inconsistent are the least likely to take up commitment (44%)

- Subjects who are dynamically consistent are more likely to take up commitment (60%)

Suggests that those with dynamic inconsistent behavior may be the least aware of their behavior

---

#Implications

How should we measure welfare in these situations?

- Utilizing "long run" preferences without temptations?

- Utilizing "short run" preferences with temptations?

- Utilizing unambiguous preferences? (preferences that are consistent across short and long run)

---

#Experimental Setup

Two experiments used (one for findings and one for replication of findings)

---

#Experimental Setup: Chicago - High Level Summary

Subjects were not told they were apart of an experiment

20 foods were offered in the experiment for home delivery

Subjects were asked to rate each food on a 1 to 7 scale

2 week food delivery program


---

#Experimental Setup: Chicago - Details

Subjects were signed up at a table at grocery stores

Recruitment was on a rolling basis

Research assistants were in charge of setting up experiment and delivering foods

Foods were selected in consultation with store managers

Foods were assured to not be bruised or damaged

In return for participating, subjects received $25-20 gift cards

---
#Experimental Setup: Chicago - Food Selection
```{r fig.width=100, fig.height=55,echo=FALSE}
library(png)
library(grid)
img <- readPNG("/Users/Brock/Desktop/Second-Year-PhD/3rd-Quarter/Experimental-Economics/Paper-Presentation/Exp-Paper-Prez/chicagofoods.png")
grid.raster(img)
```
---

#Experimental Setup: Chicago - Timeline

Week 1: Decide foods you want to receive for week 2

Week 2: Deliver foods and surprise subjects with immediate foods to exchange for

Subjects decide upon foods for week 3

Subjects were offered the choice for whether they want to make exchanges for the following week

---
#Experimental Setup: Chicago - Week 1

Subjects received a brochure and selected foods listed in the brochure as seen before

Prices for each selection were as close to a dollar as possible

Subjects then picked their preferred day and time of delivery

This elicits initial preference parameters from subjects

---
#Experimental Setup: Chicago - Week 2

Called subjects prior to arrival to ensure delivery

Upon delivery, subjects were surprised with immediate foods (their highest ranked foods) and were offered up to 4 exchanges

Immediate foods: highest rated fruit/veggie in bundle, highest rated fruit/veggie not in bundle, highest rated sweet/salty snack in bundle, highest rated sweet/salty not in bundle

After making exchanges, subjects used a new order sheet for Week 3

Asked subjects if they would like the option for exchange goods with the Week 3 delivery

---
#Experimental Setup: Chicago - Week 3

Subjects received their delivery

Subjects completed a survey and received compensation for participating

---
#Experimental Setup: Los Angeles - Details

The authors replicated the experiment in LA for validation





---
#Experimental Setup: Los Angeles - Food Selection
```{r fig.width=100, fig.height=55,echo=FALSE}
library(png)
library(grid)
img <- readPNG("/Users/Brock/Desktop/Second-Year-PhD/3rd-Quarter/Experimental-Economics/Paper-Presentation/Exp-Paper-Prez/lafoods.png")
 grid.raster(img)
```

---
#Experimental Setup: Los Angeles - Timeline

Same process

Randomly assigned half of individuals to commitment offers in second and third delivery

Other half received second surprise exchange and only had commitment offer for third visit

All 20 foods were available for exchange


---
#Timeline Summary
```{r fig.width=100, fig.height=50,echo=FALSE}
library(png)
library(grid)
img <- readPNG("/Users/Brock/Desktop/Second-Year-PhD/3rd-Quarter/Experimental-Economics/Paper-Presentation/Exp-Paper-Prez/timeline-summary.png")
 grid.raster(img)
```

---
#Design Considerations - Perishable Foods

In the Chicago study, all of the fruit and veggies were perishable so it could be an issue that subjects are exchanging not due to inconsistency, but because the fruits and veggies were bad and the sweet salty snacks were not perishable

LA made some of the sweet and salty snacks perishable as well as some fruit and veggies not perishable

---

#Design Considerations - Exchanges

In LA, all foods were exchangable (individuals could trade for any of the 20 foods during delivery)

---

#Design Considerations - Phrasing

The authors reworded the script for LA for 2 reasons

Commitment may have come across as socially desirable

Exchanges may have seemed to be harder effort for delivery

---

# Structural Analysis

Using rank order discrete choice methods

Defined foods by characteristics such as macronutrients, etc

Let $V_j = x_j \beta + \epsilon_j; j = 1,2,...$ where $x$ is food characteristics

Implies probability $F(x_1,...,x_J; \beta) = \dfrac{\exp(x_j \beta)}{\sum_{i=1}^{J}\exp(x_i \beta)}$

---

# Testing for Dynamic Inconsistency

There are two rankings of food: advance rankings and immediate rankings

Testing if the rankings are the same across both periods

---

# Results - Dynamic Inconsistency




---

# Welfare Analysis

Two different rankings imply three different welfares: advance, immediate and unambiguous

Advance: calculating utility from rankings given prior to delivery

Immediate: calculating utility from rankings given at delivery

Unambiguous: calculating utility from rankings that have not changed with/without delivery

---

# Results - Welfare Analysis














