---
title: "Dynamic Inconsistency in Food Choice"
author: "Brock Wilson"
date: "4/15/2021"
output: 
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)
library(pacman)
```


---

#Introduction

Dynamic Inconsistency arises due to individuals being present biased ($\beta \delta$ discounting)

--

A simple two period example:

Procastinate vs going to the gym

--

\begin{align}
\max_{a_t} U(c) = \beta (c_t) + \delta (v_{t+1} - c_{t+1}) + \delta^2(v_{t+2})
\end{align}

--

Cost = $c_t$

Reward = $v_t$

--

$\beta$ = 2

$\delta$ = $\dfrac{3}{4}$

Action set, $a_t$, is: $(c_t, v_{t+1}) = (0, 0), (-6, 10)$ 


---

#Introduction

Procastinate = $(0,0)$

Go to the gym = $(-6,10)$

--

In period 1:

- Utility from procastinate $= \beta*0 + \delta*0 = 0$

--

- Utility from gym $= \beta * -6 + \delta*10 = -12 + 7.5 = -5.5$

--

- Chooses procastinate

--

In period 2:

- Utility from procastinate $= 0 + \delta*0 = 0$

--

- Utility from gym $= -6 + \delta*10 = -6 + 7.5 = 1.5$

--

- Chooses to go to gym

--

Cycle repeats: individuals **procastinate today** and **go to the gym tomorrow**

---
#Introduction

#### Dynamic inconsistency: Individuals have two selves; today self and tomorrow self. If they don't share the same preferences, then there could be decision problems.

<br>

--
#### Dynamic inconsistencies pose issues: individuals want to go to the gym, but never want to go when asked

#### Current research into commitment devices and inconsistent preferences

--
<br>
#### Previous literature finds positive to zero correlation between self control problems and awareness

--
#### Suggests that those with self control problems should be aware of their problem

--
#### Caveat: all of these findings comes from lab experiments (may have experimental bias)


---

#Core Question

<br>
<br>
<br>
<br>

### What is the Relationship Between Dynamically Inconsistent Behaviour and Beliefs thereof?

---

#Summary

### Field Experiment: Home Grocery Delivery program

--
<br>
### Experimenters offer exchangable foods at time of delivery

--
<br>
### Findings imply that 40% of individuals are dynamically inconsistent

--
<br>
### Individuals who are inconsistent are the least likely to use commitment device

---

#Experimental Setup

#### Two experiments used:

- #### An initial experiment in Chicago

- #### A replication experiment in LA

<br>

<br>

--

Advance bundle = bundle pre-ordered (i.e. the bundle ordered from a prior week)

--

Commitment Demand = choosing to restrict oneself to the advance bundle

--

Immediate foods = foods offered in exchange at time of delivery


---

#Experimental Setup: Chicago

#### Natural Experiment: Subjects were not told they were apart of an experiment

--

#### 218 subjects signed up at small grocery store in a low income area

--

#### Store put out advertisement for home delivery program

--

#### 20 foods were offered in the experiment for home delivery

- #### 10 foods are sweet/salty snacks

- #### 10 foods are fruits/veggies

--

#### Manager of store provided which foods should be used (Institutional Knowledge)

--

#### Subjects were asked to rate each food on a 1 to 7 scale

--

#### 3 week food delivery program

---
#Experimental Setup: Chicago - Food Selection
```{r fig.width=100, fig.height=55,echo=FALSE}
library(png)
library(grid)
img <- readPNG("/Users/Brock/Desktop/Second-Year-PhD/3rd-Quarter/Experimental-Economics/Paper-Presentation/Exp-Paper-Prez/chicagofoods.png")
grid.raster(img)
```
---

#Experimental Setup

<br>

<br>

<br>

## 10,000 Foot View

---

#Experimental Setup: Chicago - Timeline

Week 1: 

- Decide foods you want to receive for week 2

--

Week 2: 

- Deliver foods and surprise subjects with immediate foods to exchange for 

- Subjects decide upon foods for week 3

--

- Subjects were offered the choice for whether they want to make exchanges for the following week

--

Week 3: 

- Receive last delivery

---

#Experimental Setup

<br>

<br>

<br>

## 2 Feet View

---

#Experimental Setup: Chicago - Week 1

#### Subjects were signed up at a table at grocery stores

<br>

--

#### Subjects received a brochure and selected foods listed in the brochure as seen before

--

<br>

#### Prices for each selection were as close to a dollar as possible (e.g. two apples, one Snickers bar)

--

<br>

#### Subjects then picked their preferred day and time of delivery

---
#Experimental Setup: Chicago - Week 2

#### Upon delivery, subjects were surprised with immediate foods (their highest ranked foods) and were offered up to 4 exchanges

--

#### Immediate foods: 

- Highest rated fruit/veggie in bundle

- Highest rated fruit/veggie not in bundle 

- Highest rated sweet/salty snack in bundle

- Highest rated sweet/salty snack not in bundle

--

#### After making exchanges, subjects used a new order sheet for Week 3

--

#### Asked subjects if they would like the option for exchange goods with the Week 3 delivery (IMPORTANT)

---
#Experimental Setup: Chicago - Week 3

#### Subjects received their delivery

<br>

#### Subjects completed a survey and received compensation for participating ($20 gift card)

---

#Extra Details - Chicago

Recruitment was on a rolling basis

Research assistants were in charge of setting up experiment and delivering foods

Foods were selected in consultation with store managers

Foods were assured to not be bruised or damaged

---

#Experimental Setup: Los Angeles

#### The authors replicated the experiment in LA for validation

---
#Experimental Setup: Los Angeles
```{r fig.width=100, fig.height=55,echo=FALSE}
library(png)
library(grid)
img <- readPNG("/Users/Brock/Desktop/Second-Year-PhD/3rd-Quarter/Experimental-Economics/Paper-Presentation/Exp-Paper-Prez/lafoods.png")
 grid.raster(img)
```

---
#Experimental Setup: Los Angeles

Same process

Randomly assigned half of individuals to commitment offers in second and third delivery

Other half received second surprise exchange and only had commitment offer for third visit

All 20 foods were available for exchange

---
#Timeline Summary
```{r fig.width=100, fig.height=50,echo=FALSE}
library(png)
library(grid)
img <- readPNG("/Users/Brock/Desktop/Second-Year-PhD/3rd-Quarter/Experimental-Economics/Paper-Presentation/Exp-Paper-Prez/timeline-summary.png")
 grid.raster(img)
```



---
#Design Considerations

## Perishable Foods

#### Chicago

- Fruit and veggies were perishable 

- Sweet and salty snacks were not perishable

- Exchange may be due to perishability of food and not inconsistency

#### Los Angeles

- Fruits/veggies and snacks had both perishable and non perishable options

---

#Design Considerations

## Exchanges

#### Chicago

- Only 4 foods were available for exchange (individuals' favorites)

#### Los Angeles

- All foods were available for exchange (individuals could trade for any of the 20 foods during delivery)

---

#Design Considerations

## Phrasing

#### Chicago

- Commitment may have come across as socially desirable in script

- Exchanges may have seemed to be harder effort for delivery

#### Los Angeles

- Reworded script to account for these

---

# Analysis

<br>

<br>

<br>

### Methods

---

# Analysis

Three rankings of food: 

#### Advance Rankings

- Using the food characteristics of the original basket to calculate utility

#### Immediate Rankings

- Using the food characteristics of the immediate basket to calculate utility

#### Unambiguous Rankings

- Using the food characteristics of the unchanged goods in the basket to calculate utility


---

# Analysis

## Rank order discrete choice methods

<br>

Let $V_j = x_j \beta + \epsilon_j; j = 1,2,...$

<br>

$x_j$ is food j's characteristics such as macronutrients, micronutrients, etc

<br>

Implies probability $F(x_1,...,x_J; \beta) = \dfrac{\exp(x_j \beta)}{\sum_{i=1}^{J}\exp(x_i \beta)}$

---

# Results

<br>

<br>

<br>

### 

---

# Bundle Characteristics



```{r fig.width=100, fig.height=50,echo=FALSE}
library(png)
library(grid)
img <- readPNG("/Users/Brock/Desktop/Second-Year-PhD/3rd-Quarter/Experimental-Economics/Paper-Presentation/Exp-Paper-Prez/bundle-charac.png")
 grid.raster(img)
```


Note that constants sum to 10

---

# Results - Commitment


```{r fig.width=100, fig.height=50,echo=FALSE}
library(png)
library(grid)
img <- readPNG("/Users/Brock/Desktop/Second-Year-PhD/3rd-Quarter/Experimental-Economics/Paper-Presentation/Exp-Paper-Prez/dynamic-cons.png")
 grid.raster(img)
```

---
# Results - Dynamic Inconsistency


```{r fig.width=100, fig.height=50,echo=FALSE}
library(png)
library(grid)
img <- readPNG("/Users/Brock/Desktop/Second-Year-PhD/3rd-Quarter/Experimental-Economics/Paper-Presentation/Exp-Paper-Prez/dyn-reg.png")
 grid.raster(img)
```

---

# Results - Commitment


```{r fig.width=100, fig.height=50,echo=FALSE}
library(png)
library(grid)
img <- readPNG("/Users/Brock/Desktop/Second-Year-PhD/3rd-Quarter/Experimental-Economics/Paper-Presentation/Exp-Paper-Prez/advance-immediate.png")
 grid.raster(img)
```




---
# Results - Dynamic Inconsistency


```{r fig.width=100, fig.height=50,echo=FALSE}
library(png)
library(grid)
img <- readPNG("/Users/Brock/Desktop/Second-Year-PhD/3rd-Quarter/Experimental-Economics/Paper-Presentation/Exp-Paper-Prez/dyn-utility.png")
 grid.raster(img)
```

---

# Results - Welfare Analysis


```{r fig.width=100, fig.height=50,echo=FALSE}
library(png)
library(grid)
img <- readPNG("/Users/Brock/Desktop/Second-Year-PhD/3rd-Quarter/Experimental-Economics/Paper-Presentation/Exp-Paper-Prez/welfare-1.png")
 grid.raster(img)
```

---

# Results - Welfare Analysis

```{r fig.width=100, fig.height=50,echo=FALSE}
library(png)
library(grid)
img <- readPNG("/Users/Brock/Desktop/Second-Year-PhD/3rd-Quarter/Experimental-Economics/Paper-Presentation/Exp-Paper-Prez/welfare-2.png")
 grid.raster(img)
```

---


# Results - Welfare Analysis

```{r fig.width=100, fig.height=50,echo=FALSE}
library(png)
library(grid)
img <- readPNG("/Users/Brock/Desktop/Second-Year-PhD/3rd-Quarter/Experimental-Economics/Paper-Presentation/Exp-Paper-Prez/welfare-3.png")
 grid.raster(img)
```

---
# Results - Welfare Analysis

```{r fig.width=100, fig.height=50,echo=FALSE}
library(png)
library(grid)
img <- readPNG("/Users/Brock/Desktop/Second-Year-PhD/3rd-Quarter/Experimental-Economics/Paper-Presentation/Exp-Paper-Prez/welfare-4.png")
 grid.raster(img)
```

---
# Results - Robustness

Reminder:

"In this subsection, we provide additional evidence that dynamic inconsistency is a product of preferences rather than an alternate force such as the resolution of uncertainty, changing environmental factors or noise"

---
# Results - Robustness

```{r fig.width=100, fig.height=50,echo=FALSE}
library(png)
library(grid)
img <- readPNG("/Users/Brock/Desktop/Second-Year-PhD/3rd-Quarter/Experimental-Economics/Paper-Presentation/Exp-Paper-Prez/want-vs-should.png")
 grid.raster(img)
```


---

# Conclusion

#### Dynamic inconsistences are prevalent (around 40% of individuals)

--

<br>

#### Negative correlation between consistency and awareness

--

<br>

#### Inconsistency leads to approximately a 5% decrease in welfare (depends on criterion)

--

<br>

#### Both mandating and allowing commitment does not lead to a Pareto improvement; only tailored policy


--

<br>

#### It's still hard to get people to eat their veggies

---

# Extras

---

# Findings

46% of individuals are dynamically inconsistent when commitment is not available

Regularly individuals substitute fruits and veggies for caloric-dense and high fat foods

When commitment is available, 53% of individuals take it up

- Subjects who are dynamically inconsistent are the least likely to take up commitment (44%)

- Subjects who are dynamically consistent are more likely to take up commitment (60%)

Suggests that those with dynamic inconsistent behavior may be the least aware of their behavior

---

# Implications

How should we measure welfare in these situations?

- Utilizing "long run" preferences without temptations?

- Utilizing "short run" preferences with temptations?

- Utilizing unambiguous preferences? (preferences that are consistent across short and long run)

---








